---
title: "Functions"
format: html
editor: visual
---

## Function One

Makes sequence with length of your choosing

```{r}
Sequence <- function(sequence_length){
  seq <- sample(c("A", "T", "G", "C"), size = sequence_length, replace = TRUE)
  seq_clean <- paste0(seq, collapse = "")
  return(seq_clean)
}

```

## Function Two

Replaces "T" with "U" in a sequence

```{r}
substitueDNARNA <- function(DNA_sequence){
  RNA_sequence <- gsub("T", "U", DNA_sequence)
  return(RNA_sequence)}

```

## Function Three

Takes DNA sequence, indicating start of the sequence at one, and then reading it in codons, also known as the reading frame of the sequence.

```{r}
settingReadingFrameCodons <- function(dnaSequence, start = 1){
  sequenceLength <- nchar(dnaSequence)
  codons <- substring(dnaSequence,
                      first = seq(from = start, to = sequenceLength - 3 + 1, by = 3),
                      last = seq(from = 3 + start - 1, to = sequenceLength, by = 3))
  return(codons)
}

```

## Function Four

Translates RNA, generating the translated amino acids of the sequence.

```{r}
generateAAFromCodons <- function(codons){
  AA <- paste0(codon_table[codons], collapse = "")
  return(AA)
}
```

## Function Five

Plots the frequency of each amino acid observed in the sequence.

```{r}
generateAminoAcidFrequencyPlot <- function(Sequence){
  amino_acids <- Sequence |>
    stringr::str_split(pattern = stringr::boundary("character"), simplify = TRUE) |>
    as.character() |>
    unique()
  
  amino_acid_counts <- sapply(amino_acids, function(amino_acid) stringr::str_count(string = Sequence, pattern = amino_acid)) |>
    as.data.frame()
  
  colnames(amino_acid_counts) <- c("Counts")
  amino_acid_counts[["AminoAcid"]] <- rownames(amino_acid_counts)
  
  frequency_plot <- amino_acid_counts |>
    ggplot2::ggplot(ggplot2::aes(x = AminoAcid, y = Counts, fill = AminoAcid)) +
    ggplot2::geom_col() +
    ggplot2::theme_bw() +
    ggplot2::theme(legend.position = "none")
  
  return(frequency_plot)
}
```
